MSc Thesis

### 2023-05-02 manuscript/main.tex
clarified p val for mn ~ si reg in qtl ch

When the plants had reached maturity, we harvested the entire above-ground portion of each plant. For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022. We placed harvested material in labelled paper bags, and dried it a 60ºC drying oven for 48 hours. To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers. Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation. We picked leaves until approximately 200mg of dry leaf was collected. Some plants did not yield enough leaf tissue to meet to 200mg threshold. We pooled leaf material from within sites. For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample. Overall, we were left with approximately 115 useable genotypes from each site. We packed dried leaves samples were 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz. We stored the resulting leaf powder sealed until XRF analysis.

@@

To analyse the silicon and manganese content of the accessions, we followed the XRF procedure presented in \textcite{reidinger_rapid_2012}. In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand. For beam 1 (Mn), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time. Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy. For each pellet, we took two technical replicates, scanning each side of the pellet once. To minimize cross-contamination between samples, we cleaned the pellet press and XRF device after each sample. We calibrated our measurments against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).

@@

To look for evidence of root exudation driving silicon content, we tested a correlation between observed leaf silicon and manganese content. Prior to the analysis, we plotted histograms of silicon and manganese concentrations. We observed right-skew for both elements and applied a log-transformation. To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$). We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:

@@

To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}. For brevity, this methodology only describes the steps we took using the data generated from \textcite{gaurav_population_2022}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative. For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers. We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.

@@

Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis. After pooling, we were left with 359 samples across three environments. Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%. There were notable differences in silicon content based on the growing environment. The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132. Silicon and manganese showed significant positive correlation ($\beta = 0.278\pm0.051, z = 5.46, p<0.0001$) (Figure \ref{Fig:mn_si_regression}). Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}). My results for manganese content are less clear. I detected no regions that met the threshold for significance, though there were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}). Within the plants, silicon and manganese content were correlated ($R^2$ = 0.15, p = 0.049) (Figure 3).

>>>>

When the plants had reached maturity, we harvested the entire above-ground portion of each plant. For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022. We placed harvested material in labelled paper bags, and dried it a 60ºC drying oven for 48 hours. To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers. Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation. We picked leaves until approximately 200mg of dry leaf was collected. Some plants did not yield enough leaf tissue to meet to 200mg threshold. We pooled leaf material from within sites. For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample. Overall, we were left with 359 samples across three sites, making approximately 115 useable genotypes per site. We packed dried leaf samples in 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz. We stored the resulting leaf powder sealed until XRF analysis.

@@

To analyse the silicon and manganese content of the accessions, we followed the XRF procedure presented in \textcite{reidinger_rapid_2012}. In short, we pressed leaf powder into 13mm diameter pellets at 300 bar of pressure and analysed the resulting pellets in an Olympus Vanta p-XRF device mounted in a bench stand. For beam 1 (manganese), we used a 20 second read time, while for beam 2 (Si), we used a 45 second read time. Based on preliminary trials, we determined these times to be a suitable trade-off between throughput and accuracy. For each pellet, we took two technical replicates, scanning each side of the pellet once. To minimize cross-contamination between samples, we cleaned the pellet press and XRF device after each sample. We calibrated our measurments against a standard curve of methyl-cellulose spiked with silicic acid, as well as certified reference materials (WEPAL-IPE-151, WEPAL-IPE-152).

@@

To look for evidence of root exudation associated with silicon content, we tested a correlation between observed leaf silicon and manganese content. Prior to the analysis, we plotted histograms of silicon and manganese concentrations. We observed right-skew for both elements and applied a log-transformation. To control for intersite-variation we further transformed our measurements from log(concentration) to site-specific standard scores ($\frac{X - \mu}{\sigma}$). We then used the \verb|lmerTest| and \verb|MuMIn| (\cite{kuznetsova_2017_lmerTest,barton_2023_mumin}) package in R 4.2.2 (\cite{r_core_team_2022}) to implement the following model:

@@

To perform the GWAS analysis, we followed the methodology and code published in \textcite{gaurav_population_2022}. For brevity, the methodology of this manuscript only describes the steps we took using the data generated from \textcite{gaurav_population_2022}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. As per \textcite{gaurav_population_2022}, to reduce the computational intensity of my analysis, we prefiltered the total $k$-mer matrix to remove $k$-mers with a low chance of being informative. For each environment, we ran the GWAS, filtered $k$-mers with an association score of $<6$, and plotted the remaining $k$-mers. We calculated our bonferroni correction threshold using the \verb|knum_bonf.py| script provided by \textcite{gaurav_population_2022}.

@@

Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis. After pooling, we were left with 359 samples across three environments. Silicon content in \textit{Aegilops tuaschii} ranged from 0.784\% to 11.473\%. There were notable differences in silicon content based on the growing environment. The glasshouse plants averaged 1.450\% $\pm$ 0.032 (SE), while the two outdoor environments averaged 4.935\% $\pm$ 0.108 and 6.471\% $\pm$ 0.132. Silicon and manganese showed significant positive correlation ($\beta = 0.278\pm0.051, z = 5.46, p<0.0001$) (Figure \ref{Fig:mn_si_regression}). Overall, my analysis revealed no regions of the \textit{Aegilops tauschii} genome that have significant associations across environments with silicon content (Figure \ref{Fig:si_peak_plot}). For manganese, we observed numerous peaks in each environment, yet only a few peaks were consistent across environment. My results for manganese content are less clear. (Figure \ref{Fig:mn_peak_plot}). Within the plants, silicon and manganese content were correlated ($R^2$ = [How to get R2 from a lmer Model??], p = 0.0043) (Figure 3).
